Community Detection Methods for the Generalized Random Dot Product Graph Model

### 2021-05-18 proposal/slides.Rmd
updated slides

---
title: Community Detection Methods for the Generalized Random Dot Product Graph Model
subtitle: Dissertation Proposal Defense
author: John Koo
date: 'TBD'
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
---

@@

@@

* Preliminaries
  * Block Models and Community Detection
  * (Generalized) Random Dot Product Graphs
  * Connecting Block Models to the (G)RDPG
* The Popularity Adjusted Block Model
  * Connecting the PABM to the GRDPG
  * Subspace Clustering
* Community Detection for the (G)RDPG

# Preliminaries

@@

:::: {.column width=67%}

@@

* $G = (V, E)$
  * Undirected, unweighted, no self-loops
  * $|V| = n$
  * $|E| \leq n (n - 1) / 2$
* Adjacency matrix $A \in \{0, 1\}^{n \times n}$
  * $A_{ij} = \begin{cases} 
  1 & \exists \text{ edge between } i \text{ and } j \\
  0 & \text{else}
  \end{cases}$
* Edge probability matrix $P \in [0, 1]^{n \times n}$
* $A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ for $i < j$

@@

```{r, fig.height = 2, fig.width = 1.5}
n <- 2 ** 5
p <- 1 / log(n)
P <- matrix(p, nrow = n, ncol = n)
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4)
```

@@

* $A \sim BernoulliGraph(P)$
  * (Hidden) labels $z_1, ..., z_n \in [K]$
  * $P_{ij} = f(z_i, z_j, \cdot)$

* Example: Stochastic Block Model with two communities
    * $K = 2$
  * $P_{ij} = \begin{cases} 
  p & z_i = z_j = 1 \\
  q & z_i = z_j = 2 \\
  r & z_i \neq z_j
  \end{cases}$

```{r, fig.height = 1.5, fig.width = 4}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z))
```

@@

* Erdos-Renyi Model (1959)
  * $P_{ij} = \theta$
  * Not a block model
* Stochastic Block Model (Lorrain and White, 1971)
  * $P_{ij} = \theta_{z_i z_j}$
  * $K (K + 1) / 2$ parameters $\theta_{kl}$
* Degree Corrected Block Model (Karrer and Newman, 2011)
  * $P_{ij} = \theta_{z_i z_j} \omega_i \omega_j$
  * $K (K + 1) / 2 + n$ parameters $\theta_{kl}$, $\omega_i$
* Popularity Adjusted Block Model (Sengupta and Chen, 2017)
  * $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
  * $K n$ parameters $\lambda_{ik}$

@@

* PABM $\to$ DCBM: $\lambda_{ik} = \sqrt{\theta_{z_i k}} \omega_i$
* DCBM $\to$ SBM: $\omega_i = 1$
* SBM $\to$ Erdos-Renyi: $\theta_{kl} = \theta$

\begin{center}
```{r, out.width = '150px'}
knitr::include_graphics('noroozi-pensky-hierarchy.png')
```

\tiny{Majid Noroozi and Marianna Pensky. "The Hierarchy of Block Models". March 2021.}
\end{center}

* Example: DCBM ($p_{k, l, i, j} = \theta_{kl} \omega_i \omega_j$)

* NP-complete
  * Expectation-Maximization
  * Bayesian methods
  * **Spectral methods**

@@

* Random Dot Product Graph $A \sim RDPG(X)$
  * Latent vectors $x_1, ..., x_n \in \mathcal{X} \subset \mathbb{R}^d$
    * $\mathcal{X} = \{x, y : 0 \leq x^\top y \leq 1\}$
  * Data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$
  * Edge probability matrix $P = X X^\top$
  * Adjacency matrix $A \sim BernoulliGraph(P)$

* Generalized Random Dot Product Graph $A \sim GRDPG_{p, q}(X)$
  * Latent vectors $x_1, ..., x_n \in \mathcal{X} \subset \mathbb{R}^d$
    * $\mathcal{X} = \{x, y : 0 \leq x^\top I_{p, q} y \leq 1\}$
    * $I_{p, q} = blockdiag(I_p, -I_q)$
    * $p + q = d$
  * Data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$
  * Edge probability matrix $P = X I_{pq} X^\top$
  * Adjacency matrix $A \sim BernoulliGraph(P)$

* If $X_1, ..., X_n \stackrel{iid}{\sim} F$, then $(A, X) \sim RDPG(F, n)$ or 
$(A, X) \sim GRDPG_{p, q}(F, n)$

@@

* Want to estimate $X$ given $A$
  * Alternatively, recover some property of $X$ given $A$ 
    * Interpoint distances
    * Inner products
    * Angles

* Adjacency Spectral Embedding
  * Given embedding dimension $d$, 
  $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$
    * If RDPG, use $d$ greatest eigenvalues
    * If GRDPG, use $p$ most positive and $q$ most negative eigenvalues
    * $\hat{V} \in \mathbb{R}^{n \times d}$
    * $\hat{\Lambda} \in \mathbb{R}^{d \times d}$
  * RDPG: $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$
  * GRDPG: $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$

* RDPG: $\max\limits_i \|\hat{X}_i - W_n X_i \| \stackrel{a.s.}{\to} 0$
(Athreya et al.)
* GRDPG: 
$\max\limits_i \|\hat{X}_i - Q_n X_i \| \stackrel{a.s.}{\to} 0$
(Rubin-Delanchy et al.)

@@

* All $G$ with $A \sim BernoulliGraph(P)$ are RDPG (if $P$ is 
positive semidefinite) or GRDPG
  * Includes all block models

* Example: SBM

@@

```{r, out.width = '20%', fig.height = 2.5, fig.width = 2.5}
par(mar = rep(1.75, 4))
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)
plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
```

```{r, out.width = '20%', fig.height = 2.5, fig.width = 2.5}
par(mar = rep(1.75, 4))
omega <- rbeta(n, 1, 1)
dc.matrix <- omega %*% t(omega)
P.dcbm <- P * dc.matrix
dcbm.eigen <- eigen(P.dcbm)
X.dcbm <- dcbm.eigen$vectors[, 1:2] %*% diag(dcbm.eigen$values[1:2] ** .5)
plot(X.dcbm, asp = 1, col = z, xlab = NA, ylab = NA, main = 'DCBM: Rays')

@@

```{r, out.width = '33%', fig.height = 4, fig.width = 4}
par(mar = rep(1.75, 4))
Pz <- generate.P.beta(n)
P <- Pz$P
z <- Pz$clustering
X <- embedding(P)
pairs(X, col = z, asp = 1, pch = '.', main = 'PABM: Orthogonal subspaces')
```

@@

## Popularity Adjusted Block Model (Reparameterization)

@@

* w.l.o.g., organize $P$ such that each block 
$P^{(kl)} \in [0, 1]^{n_k \times n_l}$ contains edge 
probabilities between communities $k$ and $l$
* Popularity vectors $\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
are the popularity parameters of members of community $k$ to community $l$
* $\{\lambda^{(kl)}\}_K$ is the set of $K^2$ popularity vectors
* $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$
* $A \sim BernoulliGraph(P)$

@@

Theorem: $A \sim PABM(\{\lambda^{(kl)}\}_2)$ is equivalent to 
$A \sim GRDPG_{3, 1}(X U)$ for $X$ constructed from $\{\lambda^{(kl)}\}_2$ and 
$U \in \mathbb{O}(4)$

@@

**Theorem**: $A \sim PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim GRDPG_{p, q}(X U)$ such that

* $p = K (K + 1) / 2$
* $q = K (K - 1) / 2$
* $U$ is orthogonal and predetermined for each $K$
* $X$ is block diagonal and composed of $\{\lambda^{(kl)}\}_K$
* $\implies$ if $x_i^\top$ and $x_j^\top$ are two rows of $X U$ corresponding 
to different communities, $x_i^\top x_j = 0$.

@@

**Theorem**: If $P = V \Lambda V^\top$ and $B = n V V^\top$, then $B_{ij} = 0$ 
$\forall i, j$ in different communities.

Orthogonalized Spectral Clustering algorithm:

1. Compute the eigenvectors of $A$ that correspond to the $K (K+1)/2$ most 
positive and $K (K-1) / 2$ most negative eigenvalues to construct $V$.
2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
3. Construct graph $G$ using $B$ as its similarity matrix.
4. Partition $G$ into $K$ disconnected subgraphs (e.g., using edge thresholding 
or spectral clustering).
5. Map each partition to the community labels $1, ..., K$.

**Theorem**: Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity 
matrix from OSC. 
Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\{(\log n)^{4c}\}$,

@@

**Theorem**:

@@

```{r fig.height = 3, out.width = '50%'}
set.seed(1)

n1 <- 2 ** 6
n2 <- 2 ** 6
n <- n1 + n2

a <- .6

@@

```{r, fig.height = 2, fig.width = 3, out.width = '100%'}
par(mar = rep(1, 4))
plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
     main = 'Latent configuration')
```

@@

```{r, fig.height = 2, fig.width = 3, out.width = '100%'}
par(mar = rep(1, 4))
P <- data.matrix %*% t(data.matrix)
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z))
title('Sampled graph')
```

@@

```{r, fig.height = 2, fig.width = 3, out.width = '100%'}
par(mar = rep(1, 4))
A.eigen <- eigen(A)
Xhat <- A.eigen$vectors[, 1:2] %*% diag(sqrt(A.eigen$values[1:2]))
plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
```

@@

@@

@@

* E-step
  * $\gamma_{ik} = P(Z_i = k \mid \{\pi_l\}, \{\lambda_{jl}\})$
  * $\log \gamma_{ik} \propto 
  \log \pi_k + 
  \sum_{j \neq i} \sum_l \pi_{jl} (A_{ij} \log \lambda_{il} \lambda_{jk} + 
  (1 - A_{ij}) \log(1 - \lambda_{il} \lambda_{jk}))$
* M-step
  * $\pi_k = \frac{1}{n} \sum_i \gamma_{ik}$
  * $\{\lambda_{ik}\} = \arg\max_{\{\lambda_{ik}\}} E_Z[\log L]$

@@

* $Z_i \stackrel{iid}{\sim} Categorical(\pi_1, ..., \pi_K)$
* $\lambda_{ik} \stackrel{indep}{\sim} Beta(a_{ik}, b_{ik})$

>>>>

---
title: Community Detection Methods for the Generalized Random Dot Product Graph Model
subtitle: Dissertation Proposal
author: John Koo
date: 'June 2021'
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
- \setbeamertemplate{itemize items}[circle]
---

@@

# Introduction

## Community Detection for Networks

How can we cluster the nodes of a network?

Statistical inference (parametric approach):

1. Define a generative model 
$G \mid z_1, ..., z_n, \vec{\theta} \sim P(\vec{z}, \vec{\theta})$.
2. Develop a method for obtaining estimators $f(G) = \hat{z}_1, ..., \hat{z}_n$.
3. Identify asymptotic properties of estimators and prove consistency.

@@

1. Probability Models for Networks
    * Block Models and Community Structure
    * (Generalized) Random Dot Product Graphs
    * Connecting Block Models to the (G)RDPG

\vspace*{.5\baselineskip}

2. Popularity Adjusted Block Model
    * Connecting the PABM to the GRDPG
    * Subspace Clustering for Community Detection

\vspace*{.5\baselineskip}

3. Community Detection for the (G)RDPG
    * Manifold Clustering
    * Manfolds as (G)RDPG Latent Configurations

# Probability Models for Networks

@@

:::: {.column width=62.5%}

Let $G = (V, E)$ be an undirected and unweighted graph with $|V| = n$.

$G$ is described by adjacency matrix $A$ such that
$A_{ij} = \begin{cases} 
1 & \exists \text{ edge between } i \text{ and } j \\
0 & \text{else}
\end{cases}$

@@

* $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
vertices.
* $A_{ij} \stackrel{ind}{\sim} Bernoulli(P_{ij})$ for each $i < j$.
* $A_{ji} = A_{ij}$ and $A_{ii} = 0$.

Example: If $G$ is an Erdos-Renyi graph, then $P_{ij} = \theta$.

@@

:::: {.column width=37.5%}

```{r, fig.height = 2, fig.width = 2}
n <- 2 ** 5
p <- 1 / log(n)
P <- matrix(p, nrow = n, ncol = n)
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4)
```

@@

Suppose each vertex $v_1, ..., v_n$ has hidden labels $z_1, ..., z_n \in [K]$,  
and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
Then $A \sim BernoulliGraph(P)$ is a *block model*.

Example: Stochastic Block Model with two communities

:::: column

* $z_1, ..., z_n \in \{1, 2\}$
* $P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$

* To make this an assortative SBM, set $p q > r^2$.
* In this example, $p = 1/2$, $q = 1/4$, and $r = 1/8$.

:::: column

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z))
```

@@

Erdos-Renyi Model (1959)

* $P_{ij} = \theta$
* Not a block model

Stochastic Block Model (Lorrain and White, 1971)

* $P_{ij} = \theta_{z_i z_j}$
* $K (K + 1) / 2$ parameters $\theta_{kl}$

Degree Corrected Block Model (Karrer and Newman, 2011)

* $P_{ij} = \theta_{z_i z_j} \omega_i \omega_j$
* $K (K + 1) / 2 + n$ parameters $\theta_{kl}$, $\omega_i$

Popularity Adjusted Block Model (Sengupta and Chen, 2017)

* $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
* $K n$ parameters $\lambda_{ik}$

@@

PABM $\to$ DCBM: $\lambda_{ik} = \sqrt{\theta_{z_i k}} \omega_i$

DCBM $\to$ SBM: $\omega_i = 1$

SBM $\to$ Erdos-Renyi: $\theta_{kl} = \theta$

\begin{center}
```{r, out.width = '125px'}
knitr::include_graphics('noroozi-pensky-hierarchy.png')
```

\tiny{Majid Noroozi and Marianna Pensky, 2021}
\end{center}

@@

Random Dot Product Graph $A \sim RDPG(X)$  
(Young and Scheinerman, 2007)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^d$ such that 
$x_i^\top x_j \in [0, 1]$
* $A \sim BernoulliGraph(X X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

Generalized Random Dot Product Graph $A \sim GRDPG_{p, q}(X)$  
(Rubin-Delanchy, Cape, Tang, Priebe, 2020)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = blockdiag(I_p, -I_q)$
* $A \sim BernoulliGraph(X I_{p, q} X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

If latent vectors $X_1, ..., X_n \stackrel{iid}{\sim} F$, then we write 
$(A, X) \sim RDPG(F, n)$ or $(A, X) \sim GRDPG_{p, q}(F, n)$.

@@

Want to estimate $X$ given $A$, or alternatively, 
interpoint distances, inner products, or angles.

### Adjacency Spectral Embedding

To embed in $\mathbb{R}^d$,

1. Compute $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$ 
where $\hat{\Lambda} \in \mathbb{R}^{d \times d}$ and 
$\hat{V} \in \mathbb{R}^{n \times d}$.  
For RDPG, use $d$ greatest eigenvalues;  
for GRDPG, use $p$ most positive and $q$ most negative eigenvalues.

2. For RDPG, let $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$;  
for GRDPG, let $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$.

\vspace*{.5\baselineskip}

RDPG: $\max\limits_i \|\hat{X}_i - W_n X_i \| \stackrel{a.s.}{\to} 0$
(Athreya et al., 2018)  
GRDPG: 
$\max\limits_i \|\hat{X}_i - Q_n X_i \| \stackrel{a.s.}{\to} 0$
(Rubin-Delanchy et al., 2020)

@@

All $G$ with $A \sim BernoulliGraph(P)$ are RDPG (if $P$ is 
positive semidefinite) or GRDPG (includes all block models).

Example: Assortative SBM

@@

:::: column

```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
par(mar = rep(1.75, 4))
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)
plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
```

\vspace*{.5\baselineskip}

```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
par(mar = rep(1.75, 4))
omega <- rbeta(n, 1, 1)
dc.matrix <- omega %*% t(omega)
P.dcbm <- P * dc.matrix
dcbm.eigen <- eigen(P.dcbm)
X.dcbm <- dcbm.eigen$vectors[, 1:2] %*% diag(dcbm.eigen$values[1:2] ** .5)
plot(X.dcbm, asp = 1, col = z, xlab = NA, ylab = NA, main = 'DCBM: Rays')

@@

:::: column

\vspace*{0\baselineskip}

```{r, out.width = '100%', fig.height = 4, fig.width = 4}
par(mar = rep(1.75, 4))
Pz <- generate.P.beta(n)
P <- Pz$P
z <- Pz$clustering
X <- embedding(P)
pairs(X, col = z, asp = 1, pch = '.', main = 'PABM: Orthogonal subspaces')
```

@@

## Popularity Adjusted Block Model

Definition based on Noroozi, Rimal, and Pensky (2020).

@@

1. w.l.o.g., organize $P$ such that each block 
$P^{(kl)} \in [0, 1]^{n_k \times n_l}$ contains edge 
probabilities between communities $k$ and $l$.
2. Organize parameters as vectors such that 
$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
are the popularity parameters of members of community $k$ to community $l$.  
$\{\lambda^{(kl)}\}_K$ is the set of $K^2$ popularity vectors.
3. Then we can write each block of $P$ as 
$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.
4. Sample $A \sim BernoulliGraph(P)$.

@@

**Theorem** (KTT): $A \sim PABM(\{\lambda^{(kl)}\}_2)$ is equivalent to 
$A \sim GRDPG_{3, 1}(X U)$ for $X$ constructed from $\{\lambda^{(kl)}\}_2$ and 
$U \in \mathbb{O}(4)$

@@

**Theorem** (KTT): $A \sim PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim GRDPG_{p, q}(X U)$ such that

* $p = K (K + 1) / 2$
* $q = K (K - 1) / 2$
* $U$ is orthogonal and predetermined for each $K$
* $X$ is block diagonal and composed of $\{\lambda^{(kl)}\}_K$  
$\implies$ if $x_i^\top$ and $x_j^\top$ are two rows of $X U$ corresponding 
to different communities, $x_i^\top x_j = 0$.

@@

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$, 
then $B_{ij} = 0$ $\forall i, j$ in different communities.

Orthogonal Spectral Clustering algorithm:

1. Compute the eigenvectors of $A$ that correspond to the $K (K+1)/2$ most 
positive and $K (K-1) / 2$ most negative eigenvalues to construct $V$.
2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
3. Construct graph $G$ using $B$ as its similarity matrix.
4. Partition $G$ into $K$ disconnected subgraphs (e.g., using edge thresholding 
or spectral clustering). Map each partition to the community labels $1, ..., K$.

**Theorem** (KTT): 
Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC. 
Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\{(\log n)^{4c}\}$,

@@

**Theorem** (KTT):

@@

```{r fig.height = 3, out.width = '50%'}
set.seed(123456)

n1 <- 2 ** 7
n2 <- 2 ** 7
n <- n1 + n2

a <- .5

@@

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
     main = 'Latent configuration')
```

@@

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
P <- data.matrix %*% t(data.matrix)
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z))
title('RDPG')
```

@@

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
A.eigen <- eigen(A)
Xhat <- A.eigen$vectors[, 1:2] %*% diag(sqrt(A.eigen$values[1:2]))
plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
```

@@

## Parameter Estimation

@@

Example: DCBM ($p_{k, l, i, j} = \theta_{kl} \omega_i \omega_j$)

* NP-complete
  * Expectation-Maximization
  * Bayesian methods
  * Spectral methods

@@

E-step

* $\gamma_{ik} = P(Z_i = k \mid \{\pi_l\}, \{\lambda_{jl}\})$
* $\log \gamma_{ik} \propto 
\log \pi_k + 
\sum_{j \neq i} \sum_l \pi_{jl} (A_{ij} \log \lambda_{il} \lambda_{jk} + 
(1 - A_{ij}) \log(1 - \lambda_{il} \lambda_{jk}))$

M-step

* $\pi_k = \frac{1}{n} \sum_i \gamma_{ik}$
* $\{\lambda_{ik}\} = \arg\max_{\{\lambda_{ik}\}} E_Z[\log L]$

@@

* $Z_i \stackrel{iid}{\sim} Categorical(\pi_1, ..., \pi_K)$
* $\lambda_{ik} \stackrel{ind}{\sim} Beta(a_{ik}, b_{ik})$
